### 2013-06-13 paper/renderers.tex

\documentclass{article}
\usepackage{tikz}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{proof}
\usepackage{qtree}
\usepackage{amssymb,amsthm}
\usepackage[left=3cm,top=3cm,right=4cm,nohead,bottom=3cm]{geometry}
\parindent 0pt

\title{\Large\textbf{A Generative Representation of Sampling-Based Visibility Algorithms}}
\author{Nicolas Feltman et al.}
\begin{document}
\maketitle

\begin{center}
There has been considerable research activity in the field of sampling-based visibility algorithms for the past 25 years.  To achieve high performance, these algorithms have become very specialized to workloads and hardware domains.  [why is this a problem]. We present a novel representation of the space of sampling-based visibility algorithms using notions of decomposition, bounds formation, bounds testing, and map-reduction as primitives.  We also present a compiler capable of translating our representation to single-core code. 
\end{center}

\end{document}