### 2014-10-17 paper/umut.tex
taking notes

\section{Umut's notes}

...To be added to the paper...

Answer the following question for all of the related work:

What happens when we try to write quickselect as a meta program?

\newcommand{\drun}[2]{\lVert{#2}\rVert_{#1}}
\paragraph{Partial Evaluation.}

The following definitions are adapted from the partial evaluation
book.  For simplicity and uniformity, we assume that data and programs
are draws from the same set and that ill-behaved programs and
non-terminating programs all return $bottom$.

Definition[Residual].

Let $p \in P$ be a program in a language $P$, which takes two inputs
$d_s \in P$ (static data) and $d_d \in P$ (dynamic data).  The program
$r \in P$ is a {\em residual for $p$ with respect to $d_s$} if for all
$d_d$, residual behaves that same as the $p$ for static data $d_s$,
i.e., 
\[
\drun{P}{p} (d_s, d_d) = \drun{P}{r} (d_d).
\]

Definition[Partial Evaluator]
A {\em partial evaluator} or a {\em program specializer} is a program
$e \in P$ (for ``evaluate'') such that for every $p \in P$, and every
$d_s, d_d \in P$, 
\[
\drun{P}{p} (d_s, d_d) = \drun{P}{\drun{P}{e}(p, d_s)} (d_d).
\]

Foundations of partial evaluation go back to Kleene's s-n-m theorem
from 1952, which constructively proved the existence of partial
evaluators by using Turing machines~\cite{Kleene52}.  Kleene did not
intend to improve efficiency, however. The first use of the term
``partial evaluation'' appears in Lombardi and Raphael (1964), in the
context of a paper on incremental camputation. Futamura (1971)
considered self-application of partial evaluation, also using it for
the purpose of compiler generation.  Ershov enriched Futamura's work
by considering double self-application, which led to compiler
generator generators, calling also Futamura's equations between
partially evaluated programs as ``Futamura projections.''

Ershov also introduced a richer form of transformation, {\em mixed
  computation} that generalizes partial evaluation.  While in some
literature, mixed computation is characterized as similar to partial
evaluation, it is significantly more general, at least by definition.

Definition [Mixed computation]. A mixed-computation for a language $P$
consist of two algorithms $m_p, m_d \in P$ for specializing the data
and the program such that for any program $p \in P$ and any data $d
\in P$ the specialized program run with the specialized data behaves
the same as the original program run with the original data, i.e.,

\[
\drun{P}{p} (d) = \drun{P}{\drun{P}{m_p} (p,d)} (\drun{P}{m_d} (p,d)).
\]

Partial evaluation is trivially a special case of mixed computation,
because it allows a transformation of the program only on the static
data, and because it allows only a simple transformation on data (the
extraction of the dynamic data).