### 2015-02-27 paper/conclusion.tex
conclusion

\section{Conclusion}

This paper presents techniques to translate staged lambda calculus
programs to multi-pass programs, which evaluate in a sequence of
passes, each of which is a conventional program.  We present a
formalization of the techniques in a way that can be implemented in
practice.  While we have also provided an implementation that
successfully splits the examples considered, we have stopped short of
offering a full proof of correctness, which we leave for future work.
We are also interested in providing a more complete implementation and
investigating how these techniques can be integrated into existing
practical systems such as existing staged programming systems for
computer graphics.  On the more conceptual side, we plan to extend the
techniques to richer source languages, for example those that include
parallelism.