### 2015-02-27 paper/conclusion.tex
Add conclusion to paper.
One editing pass over conclusion.

\section{Conclusion}

This paper presents an algorithm for translating programs written in a
staged lambda calculus into multi-pass programs, which evaluate as two
monostage programs in a sequence of two dependent passes. We present a
formalization of the techniques in a way that can be implemented in
practice.  While we have provided an implementation of this algorithm
that successfully splits the examples considered in this paper, we
have stopped short of offering a full proof of correctness, which we
leave for future work.

In addition, we are interested in providing a more complete system
implementation as well as investigating how ideas from \lang\ and the
its splitting algorithm can be integrated into existing practical
systems such as staged programming systems for computer graphics
pipelines.  On the more conceptual side, it would be interesting to
consider the extension of our splitting techniques to richer source
languages, for example, those that include parallelism.