### 2015-02-28 paper/conclusion.tex
Editing pass on conclusion.

\section{Conclusion}

This paper provides a formal presentation of an algorithm for
statically translating programs written in a two-stage lambda calculus
into multi-pass programs, which evaluate as two monostage programs in
two dependent passes.  While we have provided an implementation of
this algorithm that successfully splits the examples considered in
this paper, we have stopped short of offering a full proof of
correctness, which we leave for future work.

In addition, we are interested in exploring the behavior of splitting
for a broader set of examples as well as investigating how ideas from
\lang\ and its splitting algorithm can be integrated into existing
practical systems such as staged programming systems for computer
graphics pipelines.  On the more conceptual side, it would be
interesting to consider the extension of our splitting techniques to
richer source languages that include mutation, parallel constructs,
and more than two stages.