### 2015-07-10 paper/abstract.tex
abstract draft

%!TEX root = paper.tex

Staged programming languages assign a stage to each expression and
evaluate each expression in its assigned stage,
and can be used to describe programs where inputs arrive at different times or rates.

In this paper we present an 
algorithm for statically {\em splitting} these mixed-staged programs 
into two dependent, unstaged programs where the outputs of the first part can be efficiently reused accorss multiple invocations of the second program.

While previous algorithms for performing this transformation 
(also called \emph{pass separation} and \emph{data specialization}) operate on simpler imperative languages,
out algorithm operates on an explicitly-two-staged, typed lambda calculus \lang, 
with a $\fut$ modality denoting computation at a later stage,
and a $\curr$ modality noting purely first stage code.
Notably, our algorithm is the first to with support for mixed-stage recursive and higher-order functions.

We have proven the dynamic correctness of our splitting algorithm with respect
to a partial-evaluation semantics, and mechanized this proof in Twelf.

We have implemented the algorithm in a prototype compiler,
and we show that the ability to split a language with recursion and higher-order features
can enabled non-trivial algorithmic transformations that asymptotically improve code efficiency by nearly a linear factor,
as well as modular expression of staged programming,