### 2015-07-10 paper/intro.tex

%!TEX root = paper.tex

\section{Introduction}
\label{sec:intro}

Consider a function $F$ which takes an $x$ and $y_1,\dots,y_m$, and computes
$f(x,y_i)$ for each $y_i$:
\begin{lstlisting}
fun @$F(x, y_1, y_2 \ldots, y_m)$@ = 
  @$f(x, y_1);  f(x, y_2); \ldots; f(x, y_m)$@
\end{lstlisting}
%
In this scenario, a multivariate function is applied many times with
one argument fixed, so it is advantageous to find the computations in
$f$ that depend on $x$ and then stage program execution so they are
performed only once at the beginning of execution.

%Programs featuring this structure are particularly important in real-time
%computer graphics; in fact, modern graphics architectures \emph{require} that
%graphics computations be organized into distinct {\em passes} which perform
%increasingly fine-grained computations (for example, per object, per region of
%the screen, and per pixel, where many computations in later passes use the
%results from a single computation in an earlier pass)\,\cite{OpenGL4Spec}.
%While performant, this paradigm of programming with explicit passes results in
%complex code where invariants must hold across different passes, and local
%changes for one pass may require changes to other passes to ``plumb'' inputs
%from all other frequencies (\TODO explain what ``frequency'' means here)\,\cite{Foley:2011}.
%In other words, requiring manual separation of passes breaks key programming
%abstractions such as composition and modularity.
%Graphics researchers therefore have suggested using explicitly staged programs
%\,\cite{Proudfoot:2001,Foley:2011,He:2014}, deferring to the compiler the
%mechanical task of separating such programs into passes.

%Programming languages researchers have studied similar problems, with
%J{\o}rring and Scherlis identifying {\em frequency reduction} and {\em
% precomputation} as mechanisms for efficiently evaluating staged
%computations\,\cite{JS86-staging}.

J{\o}rring and Scherlis classify automatic program staging
transformations, such as the one described above, as forms of {\em
  frequency reduction} or {\em
  precomputation}\,\cite{JS86-staging}. To perform frequency
reduction, one identifies and hoists computations that are performed
multiple times, in order to compute them only once. To perform
precomputation, one identifies computations that can be performed in
advance, and does so---for example, at compile time if the relevant
inputs are statically known.

One common precomputation technique is partial
evaluation\,\cite{futamura71,jones96}, which relies on dynamic
compilation to specialize functions to known argument values. Going
back to our example, if $f$ specializes to a particular $v$, written
$f_v$, such that $f(v,y) = f_v(y)$, then $F$ can be specialized to $v$
as
\begin{lstlisting}
fun @$F_v(y_1, y_2, \ldots, y_m)$@ = 
  @$f_v(y_1);  f_v(y_2); \ldots; f_v(y_m)$@.
\end{lstlisting}
eliminating the need to compute $m$ times those parts of $f(v,-)$
which do not depend on the second argument.

Closely related to partial evaluation is {\em metaprogramming}, where
known values represent program code to be executed in a later stage
\,\cite{davies96,Taha97,DP01-modal,NP05-nn}. Metaprogramming enables
fine-grained control over specialization by requiring explicit
\emph{staging annotations} that mark the stage of each expression.

%(In partial evaluation, we typically assume that a
%\emph{binding-time analysis} has automatically determined which parts
%of $f$ can be specialized.)

%This is the intent of compiler optimization techniques like loop
%hoisting and common subexpression elimination, which identify the
%computations common to all invocations of $f$, and perform them once
%at the beginning of the computation.

%While partial evaluation and metaprogramming have been studied
%extensively, frequency reduction techniques have not seen the same
%attention.

While simple forms of frequency reduction include standard compiler
optimizations such as loop hoisting and common subexpression
elimination, J{\o}rring and Scherlis proposed the more general
transformation of {\em splitting} a program into multiple passes
(called pass separation in \,\cite{JS86-staging}). In our example,
splitting transforms the function $f$ into two passes $f_1$ and $f_2$
such that $f(x,y_i) = f_2(f_1(x),y_i)$. Then we evaluate $F$ by
evaluating the first pass on $x$, and using the result $z$ to evaluate
the second pass on each $y_i$.
%
\begin{lstlisting}
fun @$F_\textrm{multipass}(x, y_1, y_2, \ldots, y_m)$@ = 
  let @$z = f_1(x)$@
  in @$f_2(z, y_1);  f_2(z, y_2); \ldots; f_2(z, y_m)$@
\end{lstlisting}
%
The key difference between splitting and partial evaluation (or
metaprogramming) is that the former can be performed without access to
the first argument $x$; $F_\textrm{multipass}$ works for any $x$,
while $F_v$ is defined only for $x=v$.  Therefore, unlike partial
evaluation, splitting is a static program transformation
(``metastatic'' in parial evaluation terms) and does not require
dynamic code generation.

% Note that unlike precomputation, frequency reduction does not require
%knowledge of arguments; however, if a particular $v$ is known, we can
%use partial evaluation to specialize $F_\textrm{multipass}$ to that
%$v$.

Prior work on partial evaluation and metaprogramming has demonstrated
automatic application of these techniques on higher order functional
languages. In contrast, automatic splitting transforms have been
limited to simple languages that do not support recursion or
first-class
functions\,\cite{knoblock96,Proudfoot:2001,Foley:2011,He:2014}. In
this paper, we extend this work by presenting a splitting algorithm
for a typed lambda calculus.

As our starting point, we consider an explicitly-two-staged, typed
lambda calculus \lang\ in the style of Davies\,\cite{davies96}, with a
$\fut$ modality denoting computation in a later stage
(\ref{sec:semantics}). We then present a splitting algorithm that
transforms \lang\ program into two passes corresponding to the stages
of computation in that program (\ref{sec:splitting}). These passes are
monostaged programs, expressed in a conventional functional language.

The most interesting case involves splitting a recursive mixed-stage
function (like $f$ in the example). This yields a first pass function
$f_1$ which computes from its first-stage argument $x$ a
\emph{boundary} data structure encoding the interstage communication
in $f(x,-)$. When $f$ recurs on $x$, this boundary structure itself
has a recursive type.  The second pass is a function $f_2$ which takes
the boundary data structure $f_1(x)$ and a second-stage argument
$y_i$, and completes the computation $f(x,y_i)$, traversing the
boundary structure in light of the particular argument $y_i$.

We have implemented this splitting algorithm
(\ref{sec:implementation}) and applied it to a number of staged
programs (\ref{sec:examples}) ranging from straight-line arithmetic
operations to recursive and higher-order functions. In the case of the
quickselect algorithm, which we discuss next, the split code is in
fact asymptotically faster over many $y_i$.