### 2015-07-11 paper/conclusion.tex
compressed figures

%!TEX root = paper.tex

\section{Conclusion}

This paper presents a splitting algorithm for \lang, 
a two-staged typed lambda calculus with support for recursion and first-class functions.
The type system of \lang\ uses three worlds \bbonem, \bbonep, and \bbtwo\ to classify code by its stage,
with the modalities $\fut$ and $\curr$ providing internalizations 
of second-stage code and ground first-stage code to the mixed world.
We present a dynamic semantics that evaluates terms in two passes
and provides an eager interpretation of {\tt next} via hoisting steps.
We prove the correctness of our splitting algorithm against the dynamic semantics,
and implement this proof in Twelf.
Finally, we discuss our implementation of \lang\ and its splitting algorithm
and analyze their behavior for a number of staged programs,
including those with higher-order and recursive functions.

Looking forward, we are interested in investigating the behavior of splitting 
in the presence of richer language features such as mutation, polymorphism,
parallel constructs, and more than two stages.
Also while we have proven the dynamic correctness of our splitting algorithm,
we do not yet have a characterization of the types of the output
or of the cost behavior of output terms.