### 2015-10-12 paper/overview.tex

%!TEX root = paper.tex

\section{Overview}
\label{sec:overview}

\begin{abstrsyn}

Suppose that we wish to perform a series of order statistics queries
on a list \texttt{l}. To this end, we can use the quickselect
algorithm~\cite{Hoare:1961}, which given a list \texttt{l} and a rank
\texttt{k}, returns the element of \texttt{l} with rank \texttt{k}.
As implemented in an ML-like language in \ref{fig:qs-unstaged},
\texttt{qs} partitions \texttt{l} using the first element as
a pivot and then recurs on one of the two resulting sides, depending on
the relationship of \texttt{k} to the size \texttt{i} of the first half, in
order to find the desired element.  
If the index is out of range, a default value of \texttt{0} is returned.
Assuming that the input is uniformly
randomly ordered (which can be achieved by pre-permuting the input), \texttt{qs}
runs in expected $O(n)$ time.
%
Using \texttt{qs}, we can perform $m$ different order statistics queries with
ranks $\mathtt{k1},\dots,\mathtt{km}$ as follows:
%
\begin{lstlisting}
(qs l k1, qs l k2, ..., qs l km)
\end{lstlisting}
%
Unfortunately, this approach requires $O(n \cdot m)$ time.

We can attempt to improve on this algorithm by factoring out computations shared
between these calls to \texttt{qs}. In particular, the comparisons performed by
\texttt{partition} depend only on \texttt{l}, although different calls to
\texttt{partition} are made by each call to \texttt{qs}.

A first cut at a solution might be to first sort \texttt{l} into
\texttt{s} and then perform lookups in \texttt{s}.
Unfortunately, lookups in \texttt{s} are also $O(n)$ on average, since
\texttt{s} is a list, leading to no improvement in efficiency.

A more sophisticated solution is to construct a binary search
tree out of \texttt{l} and then simply look for the \texttt{k}th
leftmost element of that tree. To do this efficiently requires one more
innovation---augmenting the tree by storing at each node the size of its left
subtree. This approach has (expected) runtime $\Theta(n\log{n} + m\log{n})$, a
nearly linear-time improvement in $m$.

In both cases, we replace our simple solution with a two-part algorithm which
first processes the input into a lookup data structure, and then subsequently
performs fast lookups. Improving algorithms in this way is highly non-trivial,
as it requires reasoning about intricate algorithmic concerns, and implementing
more complex algorithms and data structures.

In this paper, we present a splitting algorithm which automatically transforms
staged programs into multi-part programs. In certain cases, this algorithm even
improves input algorithms in the way described above!

\subsection{Staging}

%The example itself is also implemented in our implementation.

The idea behind staged programming is to use staging annotations---in our case,
guided by types---to indicate the stage of each subterm.

In \ref{fig:qs-staged} we show a staged version of \texttt{qs}, called
\texttt{qss}, where first-stage code is colored red, and second-stage blue. In
\texttt{qss}, we regard the input list \texttt{l} as arriving in the first stage
(with type $\curr\mathrm{list}$, a list ``now''), the input rank \texttt{k} as
arriving in the second stage (with type $\fut\rmint$, an integer in the
``future''), and the result as being produced in the second stage (with type
$\fut\rmint$).

\texttt{qss} is obtained from \texttt{qs} by wrapping certain computations with
\texttt{prev} and \texttt{next}, signaling transitions between first- and
second-stage code, and $\texttt{gr}$, signaling ground terms.  
%
We also use a function
\lstinline{1`hold : ^int -> $`2`int`},
to promote first-stage integers to second-stage integers.
%
Our type system ensures that the staging annotations in \texttt{qss} are
consistent, in the sense that computations marked as first-stage cannot depend
on ones marked as second-stage.

The process of automatically adding staging annotations to unstaged code,
called \emph{binding time analysis}, has been
the subject of extensive research (\ref{sec:related}). In this paper, we do not
consider this problem, instead assuming that the annotations already exist.
In the case of \texttt{qss}, we have specifically chosen annotations which
maximize the work performed in the first stage.

\subsection{Splitting Staged Programs}

\input{figures/quickselect-split}

In the rest of this section, we present a high-level overview of the main ideas
behind our splitting algorithm, applied to \texttt{qss}.
%
Splitting \texttt{qss} yields a two-part program that creates a
probabilistically balanced, augmented binary search tree (which is isomorphic to
a treap data structure~\cite{SeidelAr96}) as an intermediate data structure. In
particular, its first part (\texttt{qs1}) constructs such a binary search tree
and its second part (\texttt{qs2}) traverses the tree, using the embedded size
information to find the element of the desired rank.
%
The code for \texttt{qs1} and \texttt{qs2} is in \ref{fig:qs-split}.

%Our splitting algorithm is guided by the local structure of the input program.
%In particular, the algorithm has no special knowledge of the quickselect
%algorithm, binary search trees, or how to perform lookups on binary search
%trees, but it is able to derive them from the input program.

Our splitting algorithm scans \texttt{qss} for first-stage computations,
gathering them into \texttt{qs1}. Given \texttt{l}, this function performs these
computations and places the information needed by the subsequent function into a
boundary data structure.
%
In particular, \texttt{qs1} performs all recursive calls and evaluates all
instances of \texttt{partition} (since it depends only on \texttt{l}). It
produces a boundary data structure that collects the results from these
recursive calls, tagged by the branch (\texttt{LT}, \texttt{EQ}, or \texttt{GT})
in which that call occurred. Since the recursive calls occur in two different
branches (\texttt{LT} and \texttt{GT}) the boundary structure is a binary tree.  
%
Lastly, it records \texttt{i} (the size of the left subtree) and \texttt{\#1 ht}
(the pivot/head of the list) in the boundary structure, because those
computations are \emph{held} for use in the second stage.
%
The final result is a binary search tree augmented with size information, and
whose keys are the pivots.

Our splitting algorithm simultaneously scans \texttt{qss} for second-stage
computations, gathering them into \texttt{qs2}. This function is given the boundary
data structure and the rank \texttt{k}, and it finishes the computation. Now that
\texttt{k} is known, the conditional on \texttt{compare k i} can be evaluated,
choosing which recursive call of \texttt{qss} is actually relevant for this
\texttt{k}. Since the boundary data structure contains the first-stage data for
\emph{all} of the recursive calls, performing these comparisons essentially
walks the tree, using the rank along with the size data \texttt{i} in order to
look up the \texttt{k}th leftmost node in the tree.

% \begin{theorem}
%   Consider an execution of \texttt{qs1} with a randomly permuted input
%   of $n$ keys and performing $m$ executions of \texttt{qs2} with the
%   result of \texttt{qs1}.  The total run-time for this computation is
%   $\Theta(n\log{n} + m\log{n})$ in expectation.
% \end{theorem}
% \begin{proof}
%   We note first that \texttt{qs1} is
%   isomorphic to a quicksort algorithm but builds a binary search tree
%   instead of combining the returned results in a list.  The run-time
%   for \texttt{qs1} is thus (expected) $\Theta(n\log{n})$.
% %
%   The depth of the tree generated by \texttt{qs1} is isomorphic to the
%     depth of a run of the quickselect algorithm, or equivalently to the
%     span of a parallel implementation of the quicksort algorithm, both
%     of which are known to be (expected) $\Theta(\log{n})$.
% %
%     By inspection of \texttt{qs2}, we can see that the algorithm takes
%     time proportional to the depth of the binary search tree, which is
%     $\Theta(\log{n})$. Consider now a random variable for each of the
%     $m$ invocations of \texttt{qs2}, each of which takes expected
%     $\Theta(\log{n})$ time.  Since the total time is the sum of these
%     random variables, and the expectation of the sum is the sum of
%     the expectations of the random variables, the bound follows.
% \end{proof}

\end{abstrsyn}